\chapter{EXP2FLUX algorithms}
\setcounter{algocf}{0}
\renewcommand*{\algorithmcfname}{EXP2FLUX algorithm}
\begin{algorithm}
\SetKwInOut{Input}{input}
\SetKwInOut{Output}{output}
\SetKwFunction{optimize}{optimize}
\SetKwFunction{fluxDistribution}{fluxDistribution}
\Input{\begin{description}
\item model1: A valid model for the `sybil' package.
\item model2: A valid model for the `sybil' package. Must have the same reactions (reaction number and reaction identifiers) as ``model1'' with different restrictions.
\item foldReport: A threshold value to be reported. All reactions with a greater or equal fold change than the given threshold are reported.
\end{description}}
\Output{The calculated fold change for the fluxes of two given metabolic models.}
\BlankLine
fluxModel1 $\leftarrow$ extract \fluxDistribution{\optimize{model1}}\;
fluxModel2 $\leftarrow$ extract \fluxDistribution{\optimize{model2}}\;
\ForEach{$\text{reaction} \in (model1 \cap model2)$}{
\eIf{fluxModel1$_{reaction}$ is equal to 0}{\KwRet fluxModel2$_{reaction}$}{\If{(1-(fluxModel2$_{reaction}$/fluxModel1$_{reaction}$ $\leq$ \texttt{foldReport)}}{\KwRet 1-(fluxModel2$_{reaction}$/fluxModel1$_{reaction}$)}}
}
\caption{fluxDifferences}
\end{algorithm}
%%%%
%%%%
\begin{algorithm}
\SetKwInOut{Input}{input}
\SetKwInOut{Output}{output}
\Input{\begin{description}
\item model: A valid model for the \texttt{`sybil'} package.
\item expression: A valid \texttt{ExpressionSet} object (one by treatment).
\item organism: A valid organism identifier for the KEGG database. List of valid organism identifiers are available in: http://rest.kegg.jp/list/organism.
\item typeID: A string to define the type of ID used in GPR's. One of \texttt{entrez} or \texttt{kegg} must be given
\item missing: A character string specifying the value to be used in missing cases; must be one of \texttt{min}, \texttt{1q}, \texttt{mean}, \texttt{median}, \texttt{3q}, or \texttt{max}.
\item scale: A boolean value to specify if data must be scaled to assign a value of 1000 as max.
\end{description}}
\Output{}
\BlankLine
\If{organism and typeID is given}{Download from the KEGG database the metabolic pathways associated to each gene}
\ForEach{non exchange reaction $\in$ model}{
\eIf{GPR is given}{
split associated GPR by \texttt{`or'} conectors to extract complex\;
\ForEach{complex}{split complex by the \texttt{`and'} conector to extract associated genes\;
\ForEach{gene $\in$ complex}{
\eIf{gene $\in$ ExpressionSet}{extract associated expression value from the ExpressionSet\;
\KwRet expression value for selected gene}{\eIf{organism and typeID is given}{
identify associated pathways to gene\;
select main pathway (more genes present in the model)\;
extract associated expression values to main pathway\;
compute the selected metric for \texttt{missing} to all expression data associated to the selected pathway}{\KwRet 0}
}}
find the minimal expression value of all genes associated in the complex\;
\KwRet identified min value}
compute sum the minimal expression values associated to all genes included in the complex that conform the GPR\;
\KwRet computed sum
}{compute the selected metric for \texttt{missing} to all expression data included in the ExpressionSet\;
\KwRet computed value}
replace \texttt{lower.bound} with computed value multiplied by -1\;
replace \texttt{upper.bound} with computed value
}
\KwRet the model with \texttt{lower} and \texttt{upper} bound modified
\caption{exp2flux}
\end{algorithm}
%%%%
%%%%